---
title: "PA1_template.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading the data
```{r}
unzip("activity.zip", exdir = "data")

activity <- read.csv("data/activity.csv")
```
##1. What is men total number of steps taken per day?
```{r}
totalStepsDay <- aggregate(steps ~ date, activity, sum)
summary(totalStepsDay)

barplot(height = totalStepsDay$steps, 
        xlab = "Date", 
        ylab = "Total steps", 
        names.arg = totalStepsDay$date,
        col = "blue", 
        main = "Total steps by date")
```
```{r}
## Calculate and report the mean and median total number of steps taken per day
medianPerDay <- median(totalStepsDay$steps, na.rm = TRUE)
meanPerDay <- mean(totalStepsDay$steps, na.rm = TRUE)
```

The mean is: `r meanPerDay` and the median is: `r medianPerDay`.

##2. What is the average daily activity pattern?
```{r}
library(dplyr)
dailyPattern <- activity %>% group_by(interval) %>% summarise(average=mean(steps, na.rm=TRUE))
plot(x=1:nrow(dailyPattern), y=dailyPattern$average, type="l", col="blue", xlab = "Interval", ylab="Average")

maxInterval <- filter(dailyPattern, average==max(average))
```

Interval `r maxInterval$interval` contains the maximum number of steps `r maxInterval$average` steps.

##3. Imputing Missing Values
### 1. Calculating and report the total number of missing values in the dataset.
```{r}
naNumber <- sum(is.na(activity$steps))
```

The total number of missing values are `r naNumber`.

### 2. Devise a astrategy for filling in all of the missing values.
There are a lot of day without values, so they will be filled in with the mean value for the 5 minute interval.

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
filledIn <- numeric(nrow(activity))
for (i in 1:nrow(activity)) {
        if (is.na(activity[i,"steps"]) == TRUE) {
                filledIn[i] <- dailyPattern[i,"average"]
        } else {
                filledIn[i] <- activity[i,"steps"]
        }
}
activityFilledIn <- mutate(activity,stepsFilledIn=filledIn)

```

### 4. Make a histogram of the total number of steps taken each day and calclate and report the mean and median total number of steps taken per day.
```{r}
totalStepsDayFilledIn <- aggregate(steps ~ date, activityFilledIn, sum)
summary(totalStepsDayFilledIn)

barplot(height = totalStepsDayFilledIn$steps, 
        xlab = "Date", 
        ylab = "Total steps", 
        names.arg = totalStepsDayFilledIn$date,
        col = "blue", 
        main = "Total steps by date filled in")

medianPerDayFilledIn <- median(totalStepsDayFilledIn$steps, na.rm = TRUE)
meanPerDayFilledIn <- mean(totalStepsDayFilledIn$steps, na.rm = TRUE)
```

The mean is: `r meanPerDayFilledIn` and the median is: `r medianPerDayFilledIn`.

There were none significant difference between the original data set and the filled data set.

### 5. Are there differences in activity patterns between weekdays and weekends?.
```{r}
library(ggplot2)
dayW <- weekdays(as.Date(activityFilledIn$date))
for (i in 1:nrow(activityFilledIn)) {
        if (dayW[i] %in% c("domingo", "sábado")) {
                dayW[i] <- "Weekend"
        } else {
                dayW[i] <- "Weekday"
        }
} 

activityFilledIn <- mutate(activityFilledIn,weekday=dayW)

dailyPatternFilledIn <- activityFilledIn %>% group_by(interval, weekday) %>% summarise(average=mean(steps, na.rm=TRUE))
qplot(x=interval, y=average, data=dailyPatternFilledIn, geom="line", facets=weekday~.)

```
